What paradigms can webcam eye-tracking be used for? Attempted replications of five ``classic'' cognitive science experiments

### 2024-11-17 group-a/E1-analysis-comparison-to-Hartshorne.Rmd
Added list of figure titles and captions; edited some captions.

```{r E1b-calib-effect, fig.cap='Calibration scores in Exp. 1b, plotted against the verb effect, operationalized as the latency of the first target fixation in the non-restricting verb condition minus that in the restricting verb condition.', out.width="50%", fig.align="center"}
#I picked this one because it's the same as in E1a
ggplot(jh.cali.analy.3, aes(x=M, y=diff))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Verb Effect (ms)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank()) +
  theme(panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      
    axis.text = element_text(size=14))

@@

```{r E1b-spaghetti-fig, fig.cap = "Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. The vertical line marks the mean noun onset time across trials and conditions.", out.width="75%", fig.align="center"}

>>>>

```{r E1b-calib-effect, fig.cap='Calibration scores in Exp. 1b, plotted against the verb effect. This was operationalized as the latency of the first target fixation in the non-restricting verb condition minus that in the restricting verb condition.', out.width="50%", fig.align="center"}
#I picked this one because it's the same as in E1a
ggplot(jh.cali.analy.3, aes(x=M, y=diff))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Verb Effect (ms)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank()) +
  theme(panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      
    axis.text = element_text(size=14))

@@

```{r E1b-spaghetti-fig, fig.cap = "Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. Results from our web-based study (Panel A; Experiment 1a) are contrasted with the current results (Panel B). The vertical line marks the mean noun onset time across trials and conditions.", out.width="75%", fig.align="center"}

### 2024-11-17 group-a/E1-analysis-replication.Rmd
Added list of figure titles and captions; edited some captions.

```{r E1-calib-effect, fig.cap='Calibration scores plotted against the verb effect on first fixations (the latency of the first target fixation in the non-restricting verb condition minus that in the restricting verb condition).', out.width="50%", fig.align="center"}
#I picked this one because the DV is easiest to explain, and because it's the DV where the effect comes out once you drop the "bad" participants
ggplot(cali.analy.3, aes(x=M, y=diff))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Verb Effect (ms)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      
    axis.text = element_text(size=14))

@@

```{r E1a-spaghetti-fig, fig.cap = "Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. The vertical line marks the mean noun onset time across trials and conditions.", out.width="75%", fig.align="center"}

>>>>

```{r E1-calib-effect, fig.cap='Calibration scores plotted against the verb effect on first fixations. This was operationalized as the latency of the first target fixation in the non-restricting verb condition minus that in the restricting verb condition.', out.width="50%", fig.align="center"}
#I picked this one because the DV is easiest to explain, and because it's the DV where the effect comes out once you drop the "bad" participants
ggplot(cali.analy.3, aes(x=M, y=diff))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Verb Effect (ms)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      
    axis.text = element_text(size=14))

@@

```{r E1a-spaghetti-fig, fig.cap = "Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. Results from an in-lab study with an eyetracker (Panel A; @james2023language) are contrasted with the current results (Panel B). The vertical line marks the mean noun onset time across trials and conditions.", out.width="75%", fig.align="center"}

### 2024-11-17 group-b/E2-analysis-calibration.Rmd
Added list of figure titles and captions; edited some captions.

```{r E2-calib-effect, fig.cap = 'Calibration scores plotted against gaze bias (the difference between the proportion of looks to the critical quadrant minus the average proportion of looks to the average of the other three quadrants).'}
ggplot(eyetracking.effects.by.subj, aes(x = mean_percent_in_roi, y = crit_bias))+
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x) +
  labs(x="Calibration Score", y="Gaze Bias")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

>>>>

```{r E2-calib-effect, fig.cap = 'Calibration scores plotted against gaze bias. This was operationalized as the difference between the proportion of looks to the critical quadrant minus the average proportion of looks to the average of the other three quadrants.', out.width="50%", fig.align="center"}
ggplot(eyetracking.effects.by.subj, aes(x = mean_percent_in_roi, y = crit_bias))+
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x) +
  labs(x="Calibration Score", y="Gaze Bias")+
  theme_bw()+
  theme(panel.grid.minor = element_blank()) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

### 2024-11-17 group-b/E2-analysis-replication.Rmd
Added list of figure titles and captions; edited some captions.

```{r E2-gaze-fig-fixed, fig.cap = "Proportion of eye-gaze to critical quadrant and other three quadrants durign memory retrieval.", eval = F, out.width="50%", fig.align="center"}
fig.fixed<-ggplot(fixed.view.summary.condition1.data, aes(x=normalized_quadrant, y=M, ymax=M+SE, ymin=M-SE, fill = normalized_quadrant == "critical" )) +
  geom_col(position=position_dodge(), color = "black")+
  geom_errorbar(position=position_dodge(width=0.9), width=0.1)+
  scale_fill_brewer(palette="Set1",)+
  theme_classic()+
  labs(x = "Quadrant", y = "Mean proportion of looks")+
  theme(legend.position = "none")    theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14), 
    axis.text = element_text(size=14)
  )

@@

```{r E2-rt-acc-fig, fig.cap = "Accuracy and response times during memory retrieval.", out.width="50%", fig.align="center"}
acc_rt_fig<-acc_fig + rt_fig + plot_layout(guides = "collect")
acc_rt_fig
```

>>>>

```{r E2-gaze-fig-fixed, fig.cap = "Proportion of eye-gaze to critical quadrant and other three quadrants durign memory retrieval in Experiment 2.", eval = F, out.width="50%", fig.align="center"}
fig.fixed<-ggplot(fixed.view.summary.condition1.data, aes(x=normalized_quadrant, y=M, ymax=M+SE, ymin=M-SE, fill = normalized_quadrant == "critical" )) +
  geom_col(position=position_dodge(), color = "black")+
  geom_errorbar(position=position_dodge(width=0.9), width=0.1)+
  scale_fill_brewer(palette="Set1",)+
  theme_classic()+
  labs(x = "Quadrant", y = "Mean proportion of looks")+
  theme(legend.position = "none")    theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14), 
    axis.text = element_text(size=14)
  )

@@

```{r E2-rt-acc-fig, fig.cap = "Accuracy and response times during memory retrieval by statement type and viewing condition in Experiment 2.", out.width="50%", fig.align="center"}
acc_rt_fig<-acc_fig + rt_fig + plot_layout(guides = "collect")
acc_rt_fig
```

### 2024-11-17 group-b/E2-calibration-figure.Rmd
Added list of figure titles and captions; edited some captions.

```{r E2-calibration-figure, echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 2. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.(In this experiment all the same locations were used for both calibration and validation.)", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```

>>>>

```{r E2-calibration-figure, echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 2. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration. In this experiment, all the same locations were used for both calibration and validation.", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```

### 2024-11-17 group-c/E3-analysis-calibration.Rmd
Added list of figure titles and captions; edited some captions.

To see if calibration success is correlated with the eye tracking effects, we calculated a calibration score for each participant. The calibration score was the average proportion of samples within 200 pixels of the validation points during the final validation phase before the eye tracking is performed. Calibration scores were not correlated with proportion looks to novel, regardless of whether scores were computed using ROIs (see Figure\ \@ref(fig:E3-cal-Plot-looking-score-roi-by-calibration)) or split-halves (see Figure\ \@ref(fig:E3-cal-Plot-looking-score-halves-by-calibration).

```{r E3-cal-Plot-looking-score-halves-by-calibration, echo=FALSE, warning=FALSE, fig.cap = 'Calibration scores plotted against proportion looks to novel using the left-vs.-right halves coding method (proportion of gaze samples to the half of the screen containing the new image out of all gaze samples).', out.width="50%", fig.align="center"}
ggplot(merged.eyetracking, aes(x=mean_percent_in_roi, y=looking.score.M.half))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Looks-to-novel halves")+
  theme_bw()+
  theme(panel.grid.minor = element_blank()) + theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

@@

```{r E3-cal-Plot-looking-score-roi-by-calibration, echo=FALSE, warning=FALSE, fig.cap = 'Calibration scores plotted against proportion looks to novel using the ROI coding method (proportion of gaze samples to the new image out of all gaze samples).', out.width="50%", fig.align="center"}
ggplot(merged.eyetracking, aes(x=mean_percent_in_roi, y=looking.score.M.roi))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Looks-to-novel (ROIs)") +
  theme_bw()+
  theme(panel.grid.minor = element_blank()) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

>>>>

To see if calibration success is correlated with the eye tracking effects, we calculated a calibration score for each participant. The calibration score was the average proportion of samples within 200 pixels of the validation points during the final validation phase before the eye tracking is performed. Calibration scores were not correlated with proportion looks to novel, regardless of whether scores were computed using ROIs (see Figure\ \@ref(fig:E3-cal-Plot-looking-score-roi-by-calibration)) or split-halves (see Figure\ \@ref(fig:E3-cal-Plot-looking-score-halves-by-calibration)).

```{r E3-cal-Plot-looking-score-roi-by-calibration, echo=FALSE, warning=FALSE, fig.cap = 'Experiment 3 calibration scores plotted against proportion looks-to-novel using the ROI coding method. This was operationalized as proportion of gaze samples to the new image out of all gaze samples).', out.width="50%", fig.align="center"}
ggplot(merged.eyetracking, aes(x=mean_percent_in_roi, y=looking.score.M.roi))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Looks-to-novel (ROIs)") +
  theme_bw()+
  theme(panel.grid.minor = element_blank()) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

@@

```{r E3-cal-Plot-looking-score-halves-by-calibration, echo=FALSE, warning=FALSE, fig.cap = 'Experiment 3 calibration scores plotted against proportion looks-to-novel using the left-vs.-right halves coding method (proportion of gaze samples to the half of the screen containing the new image out of all gaze samples).', out.width="50%", fig.align="center"}
ggplot(merged.eyetracking, aes(x=mean_percent_in_roi, y=looking.score.M.half))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Looks-to-novel halves")+
  theme_bw()+
  theme(panel.grid.minor = element_blank()) + theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

### 2024-11-17 group-d/E4-analysis-ROIs.Rmd
Added list of figure titles and captions; edited some captions.

```{r E4-gaze-timecourse-fig-quadrants, fig.cap = "Timecourse of eye-gaze to target animal and target instrument by verb bias condition with gaze categorized based on which quadrant of the screen the coordinates fall in (as opposed to a bounding box around the image). Vertical lines indicate average onsets of animal and instrument offset by 200ms.", out.width="50%", fig.align="center"}

>>>>

```{r E4-gaze-timecourse-fig-quadrants, fig.cap = "Timecourse of eye-gaze to target animal and target instrument by verb bias condition. Gaze was categorized based on which quadrant of the screen the coordinates fall in (as opposed to a bounding box around the image). Vertical lines indicate average onsets of animal and instrument offset by 200ms.", out.width="50%", fig.align="center"}

### 2024-11-17 group-d/E4-analysis-replication.Rmd
Added list of figure titles and captions; edited some captions.

Gaze fixations were time-locked to the auditory stimulus on a trial by trial basis and categorized as being directed towards one of the four items in the display if the x, y coordinates fell within a rectangle containing the image. In order to assess how verb bias impacted sentence disambiguation as the sentence unfolded, the proportion of fixations was computed in three time windows: the verb-to-animal window (from verb onset + 200 ms to animal onset + 200 ms), the animal-to-instrument window (from animal onset + 200 ms to instrument onset + 200 ms), and the post-instrument window (from instrument onset + 200 ms to instrument onset + 1500ms + 200 ms). Results were qualitatively similar to those in the original are shown in Figure \@ref(fig:E4-gaze-timecourse-fig), though proportions of fixations to the target animal were much lower in the web version of the study. This may reflect the fact that participants in the web study are less attentive and/or the quality of the webgazer eye-tracking system is lower, relative to the Eyelink 1000 which was used for the original study.  Eyegaze results are shown in more detail in Figure \@ref(fig:E4-gaze-timecourse-fig).

@@

```{r E4-proportion-fix-by-window-both, fig.cap = "Proportion of target fixations by verb bias in the original dataset (Ryskin et al., 2017) and the current data collected online. Error bars reflect bootstrapped 95% CIs over subject means", out.width="65%", fig.align="center"}

@@

```
```{r, eval  = F}
orig.data.time.window.1 <- eyetracking.window1.summary.by.trial.orig %>% 
  filter(condition != "filler") %>% 
  mutate(condition = fct_drop(condition))
contrasts(orig.data.time.window.1$condition) <- cbind(c(-2/3, 1/3, 1/3), c(0, -1/2, 1/2))

>>>>

Gaze fixations were time-locked to the auditory stimulus on a trial by trial basis and categorized as being directed towards one of the four items in the display if the x, y coordinates fell within a rectangle containing the image. In order to assess how verb bias impacted sentence disambiguation as the sentence unfolded, the proportion of fixations was computed in three time windows: the verb-to-animal window (from verb onset + 200 ms to animal onset + 200 ms), the animal-to-instrument window (from animal onset + 200 ms to instrument onset + 200 ms), and the post-instrument window (from instrument onset + 200 ms to instrument onset + 1500ms + 200 ms). Results were qualitatively similar to those in the original as shown in Figure \@ref(fig:E4-proportion-fix-by-window-both), though proportions of fixations to the target animal were much lower in the web version of the study. This may reflect the fact that participants in the web study are less attentive and/or the quality of the webgazer eye-tracking system is lower, relative to the Eyelink 1000 which was used for the original study.  Eye-gaze results are shown in more detail in Figure \@ref(fig:E4-gaze-timecourse-fig).

@@

```{r E4-proportion-fix-by-window-both, fig.cap = "Proportion of target fixations by verb bias in the original dataset (Ryskin et al., 2017) and the current data collected online. Error bars reflect bootstrapped 95% CIs over subject means.", out.width="65%", fig.align="center"}

@@

```{r, eval  = F}
orig.data.time.window.1 <- eyetracking.window1.summary.by.trial.orig %>% 
  filter(condition != "filler") %>% 
  mutate(condition = fct_drop(condition))
contrasts(orig.data.time.window.1$condition) <- cbind(c(-2/3, 1/3, 1/3), c(0, -1/2, 1/2))

### 2024-11-17 group-d/E4-calibration-figure.Rmd
Added list of figure titles and captions; edited some captions.

```{r E4-calibration-figure, echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 4. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.(In this experiment all the same locations were used for both calibration and validation.)", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```

>>>>

```{r E4-calibration-figure, echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 4. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration. In this experiment, all the same locations were used for both calibration and validation.", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```

### 2024-11-17 group-e/E5-analysis-replication.Rmd
Added list of figure titles and captions; edited some captions.

(ref:groupEMain-caption) Primary results from Exp. 5. *Top* shows the original results from Shimojo and colleagues (Figures reprinted with permission[TODO]). The attractiveness judgment along with the best-fitting sigmoid  is shown in the *top left*. Results for the roundness judgment are show in the *top right*, with the best-fitting sigmoid for the attractiveness judgment depicted in a dashed line for comparison (*top right*). (*Bottom*) shows the analogous results from the replication, with the attractiveness judgments on the *bottom left* and the roundness judgments on the *bottom right*. Again, the best-fitting sigmoid for the attractiveness judgments are plotted with a dashed line alongside the roundness results, for purposes of comparison.

```{r groupEMain, fig.show="hold", out.width="45%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:groupEMain-caption)"}
# All defaults
include_graphics("figGroupEOrigA.png")
include_graphics("figGroupEOrigB.png")
plot.A
plot.R
```

>>>>

(ref:groupEMain-caption) Primary results from Exp. 5. The attractiveness judgment along with the best-fitting sigmoid  is shown on the *left*. Results for the roundness judgment are shown on the *right*, with the best-fitting sigmoid for the attractiveness judgment depicted in a dashed line for comparison.

```{r groupEMain, fig.show="hold", out.width="45%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:groupEMain-caption)"}
# All defaults
#include_graphics("figGroupEOrigA.png")
#include_graphics("figGroupEOrigB.png")
plot.A
plot.R
```

### 2024-11-17 manuscript.Rmd
Added list of figure titles and captions; edited some captions.

authornote: |
  Partial funding for JKH's contributions was provided by NSF2436092. The authors declares no competing interests.

@@

```{r E3-design-schematic, fig.cap='Schematic of the design of Experiment 3', out.width="35%", fig.align="center"}
knitr::include_graphics("group-c/E3-example-figure.jpeg")
```

@@

#Contributions

@@

#Acknowledgements

@@

#Funding information
Funding for this project was provided by the Vassar College Department of Cognitive Science (JdL) and

@@

#Competing interests
The authors have no competing interests to declare.

@@

#Supplemental material

@@

#Data accessibility statement

#Figure titles and legends

>>>>

@@

```{r E3-design-schematic, fig.cap='Schematic of the design of Experiment 3.', out.width="35%", fig.align="center"}
knitr::include_graphics("group-c/E3-example-figure.jpeg")
```

@@

# Contributions

@@

# Funding information
Funding for data collection was provided by the Vassar College Department of Cognitive Science (JdL). Partial funding for JKH's contributions was provided by NSF2436092.

@@

# Competing interests
The authors have no competing interests to declare.

@@

# Supplemental material

@@

# Data accessibility statement

@@

# Figure titles and legends

1. **Example trial from Experiment 1.** Participants would hear a sentence (e.g., "The boy will eat the cake") and respond according to whether the sentence matched the picture.

2. **Calibration and validation point locations for Experiment 1.** Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.

3. **Calibration scores plotted against the verb effect on first fixations.** This was operationalized as the latency of the first target fixation in the non-restricting verb condition minus that in the restricting verb condition.

4. **Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time.** Results from an in-lab study with an eyetracker (Panel A; @james2023language) are contrasted with the current results (Panel B). The vertical line marks the mean noun onset time across trials and conditions.

5. **Calibration scores in Exp. 1b, plotted against the verb effect** This was operationalized as the latency of the first target fixation in the non-restricting verb condition minus that in the restricting verb condition.

6. **Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time.** Results from our web-based study (Panel A; Experiment 1a) are contrasted with the current results (Panel B). The vertical line marks the mean noun onset time across trials and conditions.

7. **Example trial from Experiment 2.**

8. **Calibration and validation point locations for Experiment 2.** Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration. In this experiment, all the same locations were used for both calibration and validation.

9. **Proportion of eye-gaze to critical quadrant and other three quadrants during memory retrieval in a) fixed and b) free viewing conditions in Experiment 2.**

10. **Accuracy and response times during memory retrieval by statement type and viewing condition in Experiment 2.**

11. **Calibration scores plotted against gaze bias.** This was operationalized as the difference between the proportion of looks to the critical quadrant minus the average proportion of looks to the average of the other three quadrants.

12. **Calibration and validation point locations for Experiment 3. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.**

13. **Schematic of the design of Experiment 3.**

14. **Correlation between proportion looks-to-novel calculated using ROIs and using screen halves.**

15. **Experiment 3 calibration scores plotted against proportion looks-to-novel using the left-vs.- right halves coding method.** This was operationalized as the proportion of gaze samples to the half of the screen containing the new image out of all gaze samples.

16. **Experiment 3 calibration scores plotted against proportion looks-to-novel using the ROI coding method.** This was operationalized as the proportion of gaze samples to the new image out of all gaze samples.

17. **The relation between calibration scores and experimental effect size, defined as the correlation between Day 1 looks-to-novel and Day 2 memory performance.** Looks-to-novel were either coded using the halves method (left panels) or ROI method (right panels); memory performance was measured using confidence ratings (top panels) or reaction time for correct recognition judgments (bottom panels).

18. **Calibration and validation point locations for Experiment 4.** Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration. In this experiment, all the same locations were used for both calibration and validation.

19. **An example of a critical trial from Experiment 4 for the sentence “Rub the butterfly with the crayon.”** The butterfly is the target animal, the panda is the distractor animal, the crayon is the target instrument, and the violin is the distractor instrument.

20. **Proportion of first mouse movements by location and verb bias in the original dataset (Ryskin et al., 2017) and the current data collected online.**

21. **Proportion of target fixations by verb bias in the original dataset (Ryskin et al., 2017) and the current data collected online.** Error bars reflect bootstrapped 95% CIs over subject means.

22. **Timecourse of eye-gaze to target animal and target instrument by verb bias condition.** Vertical lines indicate average onsets of animal and instrument offset by 200ms.

23. **Calibration scores plotted against the verb bias effect (the difference in target animal fixations between modifier and instrument conditions).**

24. **Example participant’s gaze coordinates categorized into ROIs based on a) image bounding boxes and b) screen quadrants.** Magenta points indicate looks that were not categorized into an ROI.

25. **Timecourse of eye-gaze to target animal and target instrument by verb bias condition.** Gaze was categorized based on which quadrant of the screen the coordinates fall in (as opposed to a bounding box around the image). Vertical lines indicate average onsets of animal and instrument offset by 200ms.

26. **Calibration and validation point locations for Experiment 5.** Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.

27. **An example of a critical trial from Experiment 5.** Text did not appear on each screen.

28. **Primary results from Exp. 5.** The attractiveness judgment along with the best- fitting sigmoid is shown on the left. Results for the roundness judgment are shown on the right, with the best-fitting sigmoid for the attractiveness judgment depicted in a dashed line for comparison.

29. **Correlation between calibration accuracy (x-axis) and proportion of samples fixating target (y-axis) in Exp. 5.**

30. **Revised results for Exp. 5 after removing low-calibration accuracy participants.** *Left*: Eyegaze during attractiveness judgments, along with the best-fitting sigmoid. *Right*: Eyegze during roundness judgments, along with best-fitting sigmoid (best-fitting sigmoid for attractiveness is re-plotted with a dashed line for comparison).
